---
output: github_document
---

<!-- README.md is generated from README.Rmd. Please edit that file -->

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  fig.path = "man/figures/README-",
  out.width = "100%"
)
```

# PurpleAir

<!-- badges: start -->
[![Lifecycle: experimental](https://img.shields.io/badge/lifecycle-experimental-orange.svg)](https://lifecycle.r-lib.org/articles/stages.html#experimental)
[![CRAN status](https://www.r-pkg.org/badges/version/PurpleAir)](https://CRAN.R-project.org/package=PurpleAir)
[![R-CMD-check](https://github.com/cole-brokamp/PurpleAir/actions/workflows/R-CMD-check.yaml/badge.svg)](https://github.com/cole-brokamp/PurpleAir/actions/workflows/R-CMD-check.yaml)
<!-- badges: end -->

The goal of PurpleAir is to provide read access to the [PurpleAir API](https://api.purpleair.com/) to retrieve real-time and historical data from PurpleAir sensors.

Please note that usage of this package must abide by the requirements of the PurpleAir API.  From their website:

> PurpleAir was founded on principles of openness, sharing, and community. It is to strengthen these principles and ensure the success of our entire community that we put together updated [terms of service](https://www.purpleair.com/terms), [data license](https://www.purpleair.com/license), and [data attribution](https://www.purpleair.com/attribution) requirements when using PurpleAir data and this API. Please take a moment to review them and note the attribution guide and data license agreement. If you have any questions or need more information, we have an excellent resource at https://community.purpleair.com/c/data/api/. 

## Installation

Install the development version of PurpleAir with:

``` r
pak::pak("cole-brokamp/PurpleAir")
```

## Usage

```{r}
library(PurpleAir)
```

Querying data from the PurpleAir API requires a free [PurpleAir Developer API key](https://develop.purpleair.com/sign-in?redirectURL=%2Fdashboards%2Fkeys) linked to a Google account. Functions in the package each take a `purple_air_api_key` argument or your key can be stored in an environment variable called `PURPLE_AIR_API_KEY`).  To check your key, use:

```{r}
check_api_key(Sys.getenv("PURPLE_AIR_API_KEY"))
```

Get the latest data from a single PurpleAir sensor, defined by its [sensor key](https://community.purpleair.com/t/sensor-index/4000):

```{r}
get_sensor_data(sensor_index = 175413,
                fields = c("name", "last_seen", "pm2.5_cf_1", "pm2.5_atm"))
```

Get the latest data from many PurpleAir sensors, defined by their sensor keys,

```{r}
get_sensors_data(x = as.integer(c(175257, 175413)),
                 fields = c("name", "last_seen", "pm2.5_cf_1", "pm2.5_atm"))
```

a geographic bounding box,

```{r}
sf::st_bbox(
  c(
    "xmin" = -84.82030,
    "ymin" = 39.02153,
    "xmax" = -84.25633,
    "ymax" = 39.31206
  ),
  crs = 4326
) |>
  get_sensors_data(fields = c("name"))
```

or a date from which sensors must have been modified since.

```{r}
get_sensors_data(as.POSIXct(Sys.time()) - 60, fields = "name")
```

Get historical data from a single PurpleAir sensor:

```{r}
my_history <-
  get_sensor_history(
    sensor_index = 175413,
    fields = c("pm1.0_cf_1", "pm1.0_atm", "pm2.5_cf_1", "pm2.5_atm"),
    start_timestamp = as.POSIXct("2024-07-02"),
    end_timestamp = as.POSIXct("2024-07-05")
  )

my_history
```

and plot it:

```{r "example_sensor_history_plot"}
my_history |>
  tidyr::pivot_longer(cols = tidyr::starts_with("pm"), names_to = "pollutant", values_to = "concentration") |>
  ggplot2::ggplot(ggplot2::aes(time_stamp, concentration, color = pollutant)) +
  ggplot2::geom_line()
```

By default, the PurpleAir R package retries failed API requests related to an underlying HTTP error (e.g., network is down) or a transient API error (i.e., 429, 503). Before retrying each failed request, it waits about 2 seconds. Successive failed requests result in exponentially longer waiting times (`httr2::req_retry()`). Specify the maximum number of seconds to wait (by default 45) with the environment variable `PURPLE_AIR_API_RETRY_MAX_TIME`. 
