---
output: github_document
---

<!-- README.md is generated from README.Rmd. Please edit that file -->

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  fig.path = "man/figures/README-",
  out.width = "100%"
)
```

# PurpleAir

<!-- badges: start -->
[![Lifecycle: experimental](https://img.shields.io/badge/lifecycle-experimental-orange.svg)](https://lifecycle.r-lib.org/articles/stages.html#experimental)
[![CRAN status](https://www.r-pkg.org/badges/version/PurpleAir)](https://CRAN.R-project.org/package=PurpleAir)
[![R-CMD-check](https://github.com/cole-brokamp/PurpleAir/actions/workflows/R-CMD-check.yaml/badge.svg)](https://github.com/cole-brokamp/PurpleAir/actions/workflows/R-CMD-check.yaml)
<!-- badges: end -->

The goal of PurpleAir is to provide read access to the [PurpleAir API](https://api.purpleair.com/)

## Installation

Install the development version of PurpleAir with:

``` r
pak::pak("cole-brokamp/PurpleAir")
```

Querying data from the PurpleAir API requires a free [PurpleAir Developer API key](https://develop.purpleair.com/sign-in?redirectURL=%2Fdashboards%2Fkeys) linked to a Google account.

## Usage

```{r}
library(PurpleAir)
```

Get the latest data from a single PurpleAir sensor, defined by its [sensor key](https://community.purpleair.com/t/sensor-index/4000):

```{r}
get_sensor_data(sensor_index = 175413,
                fields = c("name", "last_seen", "pm2.5_cf_1", "pm2.5_atm"))
```

Get the latest data from many PurpleAir sensors, defined by their sensor keys,

```{r}
get_sensors_data(x = as.integer(c(175257, 175413)),
                 fields = c("name", "last_seen", "pm2.5_cf_1", "pm2.5_atm"))
```

a geographic bounding box,

```{r}

```

or a date from which sensors must have been modified since.

```{r}

```
